Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/conclusions.tex
[ paper ] conclusions added.

We presented a reliable proof-reconstruction in type theory of \Metis
solutions for problems in classical propositional logic without
equality. As a product of this research, we were able to build a proof-reconstruction tool written in \Haskell. This tool generates
\Agda proof-terms of \Metis \TSTP derivations. \Agda files
delivered by the aforementioned tool are supported by the
formalization in \Agda~\cite{AgdaProp,AgdaMetis} of the \Metis rules
described in Section~\ref{ssec:emulating-inferences}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{sssec:normalization} and Section~\ref{sssec:reordering}.
We chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

@@

Summarizing, we provided for each \Metis inference rule a theorem in
type theory to emulate the rule reasoning. This approach, mainly
exposed in Section~\ref{ssec:emulating-inferences},
admits some improvements, we leave it for future work. We concern
mainly for a better coverage of the \simplify rule.
Other improvements are related with computational efficiency of
rules strongly depend on reordering tasks,
Section~\ref{sssec:reordering}.
We believe some rules like \canonicalize can get some help from an analysis done in the translator tool.

One of the main contribution of this study was
increasing the trustworthiness of the automatic provers,
in our case, \Metis.
Justifying a proof of a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.

>>>>

We presented a reliable proof-reconstruction in type theory
for the propositional fragment of the \Metis prover.

We built a proof-reconstruction tool written in \Haskell
to generate \Agda proof-terms of \Metis \TSTP derivations.
\Agda files delivered by the translator contain call to theorems from the
\Agda formalizations of \Metis reasoning~\cite{AgdaProp,AgdaMetis}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{ssec:emulating-inferences}.
We chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

@@

Besides the fact, Kansos~\cite{Kanso2012} proof-reconstruction is
similar to the approach presented in Section~\ref{ssec:workflow},
the authors aim to justify some of the inference
rules by using propositional semantics. We instead perform only a syntactic analysis of the derivations as we mentioned above.

Summarizing, we provided for each \Metis inference rule a theorem in
type theory to emulate the rule reasoning. Our approach is mainly
exposed in Section~\ref{ssec:emulating-inferences}.
We detect some improvements for our approach that we leave it for future
work.
For instance, we can investigate the consequences of removing \clausify
inference by strengthen \canonicalize rule in Thm.~\ref{thm:canonicalize}
The \simplify rule is still lacking of some coverage since it is fairly complex, many cases could be omitted.
Other improvements are related with computational efficiency of
rules strongly depend on reordering tasks,
Section~\ref{sssec:reordering}.
We believe some rules like \canonicalize can get some help from an analysis done in the translator tool.

One of the main contribution of this study was
increasing the trustworthiness of the automatic prover \Metis.
Justifying a proof by a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.
